---
title: "Week 3 Assignment"
author: "Richard Pantoliano Jr"
date: "September 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

## Problem 3
Setup:
```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
print(name)
```
#(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.
```{r}
# Remove title & initials
name_scrubbed <- str_replace(name, "[[:alpha:]]{1,3}\\.","")
name_scrubbed <- str_trim(name_scrubbed)

# Process the scrubbed name list
output_list <- list()
for(i in 1:length(name_scrubbed))
{
  splitname <- str_split(name_scrubbed[i], " ")
  if(str_count(splitname[1], ",") == 0)
        output_list[i] <- paste(splitname[1], splitname[2], sep = " ")
  else
  {
    splitname[2] <- str_replace(splitname[2],",","")
    output_list[i] <- paste(splitname[2], splitname[1], sep=" ")
  }
}

print(output_list)
```

#(b)Construct a logical vector indicating whether a character has a title (i.e., Rev.and Dr.).
This code checks for two or three character alphabetical characters, followed by a period.
```{r}
has_title <- str_detect(name, "[[:alpha:]]{2,3}\\.")
print(has_title)
```
#(c) Construct a logical vector indicating whether a character has a second name.
This code splits the original name string by the space characters.  All names that contain a second name must have more than 2 parts.  For this particular data set, all individuals with second names do not have titles 
```{r }
split_names <- c(str_split(name, " "))
has_second <- (lengths(split_names) > 2) & !has_title
print(has_second)
```

## Problem 4
Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.

#(a) [0-9]+\\\\$
This pattern matches one or more digits followed by the dollar sign character
```{r prob4a}
str1 <- "test214$555 65$  3456767$ $"
print(str_extract_all(str1, "[0-9]+\\$"))
```

#(b) \\\\b[a-z]{1,4}\\\\b
This pattern matches all words in a string constructed of lower-case letters and having a length between 1 and 4 characters
```{r prob4b}
str2 <- "The boy walked the walk."
print(str_extract_all(str2, "\\b[a-z]{1,4}\\b"))
```

#(c) .*?\\\\.txt$

This pattern matches any number of characters, ending with ".txt" and that entire string occurring at the end of the line

```{r prob4c}
str3a <- "test1.txt\n test2.txt"
str3b <- "hello3.txt"
print(str_extract_all(str3a, ".*?\\.txt$"))
print(str_extract_all(str3b, ".*?\\.txt$"))
```

#(d) \\\\d{2}/\\\\d{2}/\\\\d{4}
This pattern matches 2 digits / 2 digits / 4 digits, in the same way that the USA does the mm/dd/yyyy date format
```{r prob4d}
str4 <- "99/99/9999 1/12/1111 12/1/1234 01/01/01 09/18/2016, a/wrong/example"
print(str_extract_all(str4, "\\d{2}/\\d{2}/\\d{4}"))
```

#(e) <(.+?)>.+?</\\\\1>
This pattern matches markup tags (e.g. XML or HTML)
```{r prob4e}
str5 <- "<test>hello</test><test5>world</test5>"
print(str_extract_all(str5, "<(.+?)>.+?</\\1>"))
```